---
title: "NY Solar Resource Data"
author: "Emmanuel Messori"
date: "12/09/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The solar resource data is available on the US open data website. To retrieve it we need an access key, freely available on [demand](https://developer.nrel.gov/signup/).

In particolar we'll work with these output fields:

* The Average Direct Normal Irradiance(`avg_dni`): Direct Normal Irradiance is the amount of solar radiation a surface receives per unit area.
* The Average Tilt at Latitude (`avg_lat_tilt`): Tilt at Latitude is the amount of radiation a surface receives per unit area that doesn't arrive on a direct path from the sun.
* The Average Global Horizontal Irradiance(`avg_ghi`): Global Horizontal Irradiance is the total amount of shortwave radiation a surface horizontal to the ground receives from above.


## API connection

```{r}
api_key <-"di7U1kMqN1orvvSS2efbg8tmprjRPQU2bvPnd0bd"

```

The next step is to determine the URL and the parameters to query the API. More info on the official [documentation](https://developer.nrel.gov/docs/solar/solar-resource-v1/).

```{r}
url <- "https://developer.nrel.gov/api/solar/solar_resource/v1.json"

#NY latitude
lat <- 41
#NY longitutude
lon <- -75

params <- list(api_key = api_key, lat = lat, lon = lon)
```

## Retrieving the data

We will define a function to retrieve the information through the API:

```{r}
library(httr)

solar_data_api <- function(u = url, q = params) {
  
  response <- GET(u, query = q)
  
  print(status_code(response))
  
  print(http_type(response))
  
  if (http_error(response)) {
    print(http_status(response))
    stop("Something wrong", .call = FALSE)
  }
  
  if (http_type(response) != "application/json") {
stop("API did not return json", call. = FALSE)
}
  json_content <- content(response, "text")
  
  data <- jsonlite::fromJSON(json_content)
  
  data
}
```

Now we can retrieve the data :

```{r}
data <- solar_data_api()
```
```{r}
str(data)
```
The resulting data is a nested list of lists. The 6 main lists contain information about the API, the call and the output.


## Processing the API response

We are interested in the information contained in the nested list `outputs`. This list contains in turn three lists of data:

```{r}
str(data$output)
```

We will focus on the monthly values. Now we have to convert these lists into a dataframe.

```{r}


values <- data$outputs
avg_dni <- unlist(values$avg_dni$monthly)
avg_ghi <- unlist(values$avg_ghi$monthly)
avg_lat_tilt <- unlist(values$avg_lat_tilt$monthly)
month <- month.abb

ny_sd <- dplyr::bind_cols(month = month, avg_dni = avg_dni, avg_ghi = avg_ghi,
                 avg_lat_tilt = avg_lat_tilt)


ny_sd
```

Another approach of treating the raw data consist of  simplifying the complex list and restructuring it into the desired output :


```{r}
#1. Unlist the data 
values <- unlist(values)

#2. Transform the resulting vector into a matrix of 13 rows (12 monthly values + 1 annual value)
mat <- matrix(values, nrow = 13)

#3. Convert the matrix into a dataframe
solar_data <- as.data.frame(mat)
colnames(solar_data) <- c("avg_dni", "avg_ghi", "avg_lat_tilt")

#4. Removing the annual values:
solar_data <- solar_data[-1,]
rownames(solar_data)<-NULL


```


## Visualising the data

```{r}
library(tidyverse)

theme_set(theme_bw())

ny_sd %>% pivot_longer(cols = 2:4, names_to = "measure") %>% ggplot(aes(factor(month, levels = month.abb), value, group= measure, color = measure)) + geom_line() + geom_point() + labs(title = "New York Seasonal Solar Irradiation", x="Month", y="Value") + scale_color_manual(name="Measure Type",labels=c("Average Direct Irradiation", "Average Horizonal Irradiance", "Average Tilt at Latitude"), values=c("orange3", "orangered3", "orangered4" ))


```


